---
title: "Session 2 Problem sheet"
author: "Nikolay Petrov"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

Did the *Been there, done that* section hence **most** of the solutions were mine. Notably, part of the ggplot (using the stat_summary func) was partially not mine, although adapted, and the boxcox procedure was completely new to me.

# Importing relevant libraries

```{r results = "hide", message = FALSE}
library(tidyverse) #mandatory
library(pastecs) #for easy descriptive stats
library(ggpubr) #useful for cool plots
library(car) #for levene's test - is there a more parsimonious solution?
library(gridExtra) #for plot with subplots
library(MASS) #boxcox procedure
```

# Question 1: Reading the data

The data are provided in space separate values - weirdly enough - and the colnames are offset by 1 such that the 'id' col at the beginning is missing.

Solution 1 for importing data: `tidyverse`

```{r message=FALSE, warning=FALSE}
data <- read_delim("TheDataSet.csv", " ", col_names = c("id", names(read_delim("TheDataSet.csv", " "))),
                   skip = 1)
```

Solution 2: using base `read.csv` function.

```{r}
data <- read.csv("TheDataSet.csv", sep = " ")
```

# Question 2: Exploring the data

Checking the STRucture of the data - tricky to guess IVs and Dvs as these are design driven but: 

* DVs
  + key
  + yes_key
  + correct
  + success
  + rt
* IVs
  + text
  + is_lure
  + group
  + old

Will leave to default column specifications and change as needed.

```{r}
str(data)
```

As I will be using `group` as a grouping variable, let's mutate it to a factor.

```{r}
data <- data %>%
  mutate(group = as.factor(group))
```

And let's now output the descriptive statistics.

First, overall descriptives for `rt`:

```{r}
knitr::kable(stat.desc(data$rt))
```

Descriptive stats by `group` (levels = 0/1):

```{r}
knitr::kable(unlist(by(data$rt, data$group, stat.desc))) #maybe split column and prettify
```

# Question 3: Plotting RT as a func of group

Not the prettiest of graphs, but does the job. Possible to prettify.

## Boxplots

```{r}
ggplot(data, aes(y=rt, x=group)) +
  geom_boxplot()
```

## Barplots with errorbars

```{r}
ggplot(data, aes(y=rt, x=group)) +
  stat_summary(fun = "mean", geom = "bar", position = position_dodge()) +
  stat_summary(fun.data = "mean_se", geom = "errorbar", position = position_dodge(width = 0.4),
               width = 0.2)
```

# Question 4: Exploring assumption

## Normality plot

```{r}
ggqqplot(data$rt) #using the ggpubr func
```

## Shaprio-Wilk test

```{r}
shapiro.test(data$rt)
```
## Homogeneity of variance tests

Bartlett's test:

```{r}
bartlett.test(formula = rt ~ group, data = data)
```
Levene's test:

```{r}
leveneTest(rt ~ group, data = data)
```
Bartlett's is significant, while Levene's is not - Why?

## Transformations

Before applying any transformations, we first need to check the skewness of the data as negative skew requires inversion. See [this tutorial](https://www.datanovia.com/en/lessons/transform-data-to-normal-distribution-in-r/#transformation-methods) for more information.

In this case the skewness of the data is `r moments::skewness(data$rt)`.

Let's now plot what different transformations do to the normality of our data.

```{r}
plot1 <- ggqqplot(data$rt, title = "Original")
plot2 <- ggqqplot(log(data$rt), title = "Log transformed")
plot3 <- ggqqplot(sqrt(data$rt), title = "Square root transformation")
plot4 <- ggqqplot(1/data$rt, title = "Inverse")
grid.arrange(plot1, plot2, plot3, plot4, ncol=2)
```

The inverse transformations seems most appropriate from the graph above. To confirm, let's use the Box-Cox procedure.

```{r}
lambdaList <- boxcox(rt~group, data=data)
```

In this case the lambda value is `r (lambda <- lambdaList$x[which.max(lambdaList$y)])`. Based on the criteria given [here](https://www.isixsigma.com/tools-templates/normality/making-data-normal-using-box-cox-power-transformation/), our conclusions converge that the most effective transformation would be the inverse one.