---
title: "Session 3 Solutions Master"
author: "Team Blue"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

# Background

NP: Did the *Been there, done that* section hence **most** of the solutions were mine.

# Importing relevant libraries

Even if a library is used as a one-off, please add it here.

```{r results = "hide", message = FALSE}
library(tidyverse) #mandatory
library(corrr) #very cool function for working with correlations, see https://drsimonj.svbtle.com/exploring-correlations-in-r-with-corrr
```

# 1. Importing the Dataset

Covered in Session 2 Solutions.

```{r}
data <- read.csv("TheDataSet.csv", sep = " ")
```

# 2. Creating the pokemon dataset

First, let's set the seed to ensure the dataframe is reproducible.

```{r}
set.seed(33)
```

Create the pokemon data - using variables for the repeated parameters.

```{r}
exemplars_per_type <- 100
power_sd <- 5
prevalence_sd <- 50
pokemon_data <- data.frame(
  type        = rep(c("Salamence", "Onix", "Psychic"), each = exemplars_per_type),
  power       = c(rnorm(exemplars_per_type, mean = 20, sd = power_sd), 
                  rnorm(exemplars_per_type, mean = 50, sd = power_sd),
                  rnorm(exemplars_per_type, mean = 70, sd = power_sd)),
  prevalence  = c(rnorm(exemplars_per_type, mean = 200, sd = prevalence_sd), 
                  rnorm(exemplars_per_type, mean = 500, sd = prevalence_sd),
                  rnorm(exemplars_per_type, mean = 700, sd = prevalence_sd))
)
knitr::kable(head(pokemon_data))
```

Creating a wide version (columns instead of rows) of the dataset that will allow easy correlation matrices later.

```{r}
pokemon_data_wide <- pokemon_data %>%
  group_by(type) %>%
  mutate(row = row_number()) %>% #creating unique identifier - https://stackoverflow.com/a/58837832
  pivot_wider(names_from = type,
              values_from = c(power, prevalence)) %>%
  select(-row)
knitr::kable(head(pokemon_data_wide))
```

# 3. Plotting and quantifying the relationship between power and prevalence

```{r}
plot(x = pokemon_data$power, y = pokemon_data$prevalence, main="Scatterplot of ungrouped variables",
   xlab="Power", ylab = "Prevalence")

pairs(pokemon_data_wide[1:length(colnames(pokemon_data_wide))],
   main="Scatterplots per type")
```

The overall, ungrouped pearson correlation coefficient between power and prevalence is `r round(cor(pokemon_data$power, pokemon_data$prevalence, method = "pearson"), 3)`, while the spearman coefficient is `r round(cor(pokemon_data$power, pokemon_data$prevalence, method = "spearman"), 3)`.

Pearson correlation coefficients between power and prevalence for each pokemon type:
```{r}
knitr::kable(
  pokemon_data_wide %>%
  correlate(method = "pearson", quiet = TRUE) %>%
  shave() %>%
  fashion(decimals = 3, leading_zeros = FALSE)
)
```

Spearman correlation coefficients between power and prevalence for each pokemon type:
```{r}
knitr::kable(
  pokemon_data_wide %>%
  correlate(method = "spearman", quiet = TRUE) %>%
  shave() %>%
  fashion(decimals = 3, leading_zeros = FALSE)
)
```

# 4. Visualising and quantifying the relationship between reaction times and success rate in the dataset

Creating a summary data table for easier plotting:
```{r}
summary_data <- data %>%
  group_by(code) %>%
  summarise(mean_RT = mean(rt),
            mean_SR = mean(success),
            .groups = "drop") #not 100% sure what all effects of the .groups argument are
knitr::kable(head(summary_data))
```

```{r}
pearson_rt_sr <- round(cor(summary_data$mean_RT, summary_data$mean_SR, method = "pearson"), 3)
spearman_rt_sr <- round(cor(summary_data$mean_RT, summary_data$mean_SR, method = "spearman"), 3)
plot(x = summary_data$mean_RT, y = summary_data$mean_SR,
      main = paste("Scatterplot of reaction time and success rate,
      Pearson's r = ", pearson_rt_sr, ", Spearman's rho = ", spearman_rt_sr, sep=""),
      xlab="Reaction time", ylab = "Success rate")
```










